CALVIS: Chest, wAist and peLVIS circumference from 3D human body meshes as ground truth for deep learning

### 2019-08-06 PaperCVFAD2019/egpaper_for_review.tex
add extrema

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the hip 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves (\cite{book.compu.topo}) along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j$ as
\begin{align}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
1}^{i = n}s_i
\end{align}

@@

Next, we assemble the mesh slide length vector $\vec{\mathcal{L}}$. Starting 
from the 
top $j_t$ of the bounding box we cut the mesh into 
evenly 
$m$-meters spaced slices until we reach the bounding box bottom $j_b$. For 
every 
slice at point $j$, we compute $\mathcal{BL}(j)$.

\begin{align}
\vec{\mathcal{L}} = \mathcal{BL}(j), \, j 
\in 
\{j_t, j_t+m, \cdots, j_b\}
\end{align}

Finally, we can construct the human body \textbf{mesh signature} $\mathcal{MS}: 
\mathbb{Z}^+ \to \mathbb{R}$ of mesh $\mathcal{M}$ sliced iterative by plane 
$\boldsymbol{\pi}$ every $m$-meters by setting its value to the value of 
$\vec{\mathcal{L}}$ at 
slide index $x$

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(\mathcal{M}, \boldsymbol{\pi}, m) = \vec{\mathcal{L}}(x), \, x \in 
\{0, 1, \cdots\}
\end{align}

Figure \ref{fig:hbm_signature} shows the $\mathcal{MS}$ of two meshes (male 
and female) for 
$m=0,001$ and plane $\boldsymbol{\pi}$ parallel to the floor (with normal $(0, 
1,0)$) using the library trimesh \cite{trimesh}. The 
function resembles a rotated silhouette of the human body and exhibits several 
\textit{extrema}.

@@

By comparing neighboring values of the mesh signature $\mathcal{MS}$ as defined 
in equation \ref{eq:mesh_signature} we find several \textit{extrema}. In 
general, we use these \textit{extrema} as features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms.
	\item The local maximum $M_{hc}$ with largest value other than $M_g$ 
	immediately posterior to it represents the hip circumference.
	\item The local minimum $M_{wc}$ posterior to $M_g$ but prior to $M_{hc}$ 
	represents hip circumference.
	\item The local maximum $M_{cc}$ posterior to $M_g$ but prior to $M_{wc}$ 
	represents chest circumference.
\end{enumerate}

@@

We conduct two experiments. In the first experiment we synthesize 10 human body 
meshes. Then we apply our method to calculate chest, waist and pelvis 
circumference. We evaluate the results qualitatively. We observe that the 
measurements can indeed be used to estimate the shape of a person. The second 
experiment serves as a proof-of-concept where we input the calculated human 
dimensions to an artificial neural network. The idea is to establish the 
plausibility of our approach. After having trained the network with our data, 
we proof that the network is able to conduct this task.

>>>>

Intuitively, we would like to measure the chest circumference bellow the arms 
at the widest part of the torso and the waist circumference at the 
narrowest part after the chest but above the hips. Similarly, the hip 
circumference is measured often around the widest part of hips and buttocks. We 
can formalize this intuition by considering the cross-sectional length of the 
2D intersection curves (\cite{book.compu.topo}) along the y-axis. Moreover, we 
can 
use a plane $\boldsymbol{\pi}$ parallel to the floor to intersect the mesh. 
Since $\mathcal{M}$ is 
triangulated, the boundary of this 
intersection is a collection of segments $s_i$. Therefore, we can 
determine the boundary length by specifying a point along the y-axis $j \in 
\mathbb{R}$ as
\begin{align}
\mathcal{BL}(\mathcal{M}, \boldsymbol{\pi}, j) = \sum_{i = 
1}^{i = n}s_i
\end{align}

@@

Next, we assemble the mesh slide length vector $\vec{\mathcal{L}}$. Starting 
from the 
top $j_t$ of the bounding box we sliced iterative mesh $\mathcal{M}$ with plane 
$\boldsymbol{\pi}$ every $m$-meters until we reach the bounding box bottom 
$j_b$. For every slice at point $j$, we compute $\mathcal{BL}(j)$.

\begin{align}
\vec{\mathcal{L}}(\mathcal{M}, \boldsymbol{\pi}, m) = \mathcal{BL}(j), \, j 
\in 
\{j_t, j_t+m, \cdots, j_b\}
\end{align}

Finally, we can construct the human body \textbf{mesh signature} $\mathcal{MS}: 
\mathbb{Z}^+ \to \mathbb{R}$ that maps every slice index $k \in \{0, 1, \cdots, 
|\mathcal{L}|-1\}$ to the corresponding boundary length $\mathcal{BL}(j)$ in 
the mesh slide vector $\vec{\mathcal{L}}$.

\begin{align}\label{eq:mesh_signature}
\mathcal{MS}(k) = \mathcal{BL}(j)
\end{align}

Figure \ref{fig:hbm_signature} shows the $\mathcal{MS}$ of two meshes (male 
and female) for 
$m=0,001$ and plane $\boldsymbol{\pi}$ parallel to the floor (with normal $(0, 
1,0)$) using the library trimesh \cite{trimesh}. Note that the function as 
defined by equation \ref{eq:mesh_signature} is bounded and not continuous. It 
resembles a rotated silhouette of the human body and exhibits several 
\textit{extrema}.

@@

By comparing neighboring values of the mesh signature $\mathcal{MS}$ as defined 
in equation \ref{eq:mesh_signature} we find several \textit{extrema}. In 
general, we expect these \textit{extrema} to be adequate features to 
calculate the human dimensions. More specifically, we assume that:
\begin{enumerate}
	\item The global maximum $M_g$ represents the length of the largest path 
	around the arms. From a topographic point of view $M_g$ is the highest peak 
	of $\mathcal{MS}$.   
	\item The local maximum $M_{hc}$ with largest value other than $M_g$ 
	immediately posterior to it represents the hip circumference.
	\item The local minimum $M_{wc}$ posterior to $M_g$ but prior to $M_{hc}$ 
	represents waist circumference.
	\item from a topograhpic point of view $M_g$ is a peak from  	
	The local maximum $M_{cc}$ posterior to $M_g$ but prior to $M_{wc}$ 
	represents 
	chest circumference. If this local maximum does not exist we set $M_{cc}$ 
	to the value of the path $\mathcal{BL}(j_{cc})$ with largest value $M_{cc}$ 
	posterior 
	to 
	$M_g$ and difference at least $\delta$ to it but prior to $M_{wc}$	
	represents 
	chest circumference.
\end{enumerate}

@@

We conduct two experiments. In the first experiment we synthesize 10 human body 
meshes. Then we apply CALVIS to calculate chest, waist and pelvis 
circumference. We evaluate the results qualitatively by comparing our method 
with \cite{Dibra.2016b}. We observe that the 
measurements can indeed be used to estimate the shape of a person. The second 
experiment serves as a proof-of-concept where we input the calculated human 
dimensions to an artificial neural network. The idea is to establish the 
plausibility of our approach. After having trained the network with our data, 
we proof that the network is able to conduct this task.
